### 2017-09-14 notebook/presentation/GMSE_presentation.Rmd
More details and structure

---
title: "GMSE: a general tool for management strategy evaluation"
subtitle: "https://bradduthie.shinyapps.io/GMSE_presentation/#1"
author: "Brad Duthie"
date: "19 September 2017"
output: 
  ioslides_presentation:
    logo: Confoobio-stacked-logo.png
    css: slide.css
    widescreen: no
    fig_width: 5
    fig_height: 5
runtime: shiny
---

```{r, echo=FALSE}
library(gamesGA);
library(GMSE);
library(knitr);
```

## Introduction

## Process of adaptive management

## Management Strategy Evaluation

![](mse_overview_sm.png)

## Uncertainty

##

![](wildlife_conflict_sm.jpg)

<center>
**Image**: [Public domain](https://commons.wikimedia.org/wiki/File:Human-wildlife_conflict.jpg)
</center>

## Conservation conflict

- People with strongly held views clash over objectives
- Negative effect on sustainability, biodiversity, and human livelihood

## Background on ConFooBio

Resolving conflicts between food security and biodiversity conservation under uncertainty

<div class = "columns-2">

![](confoobio_sans.png)

- How to solve conflict
- What leads to conflict
- Impact on sustainability

All of the above are characterised by uncertainty

</div>

<!--- Get the ERC logo and EU flag in here? --->

## Game theory: conflict and cooperation

<div class="columns-2">

- Formalisation and study of conflict and cooperation between rational decision-makers
- Broadly encompasses combinatorial games (e.g., tic-tac-toe, chess) and evolutionary invasion analysis (adaptive dynamics)
- Compatibile with agent-based modelling

![](payoff_matrix_sm.png)

</div>
<hr>

[Adami et al. (2016) Phys. Life. Rev. 19:1-26](http://dx.doi.org/10.1016/j.plrev.2016.08.015)

## {.smaller}

<div class="columns-2">

![Public license](US_USSR_nukes_sm.png).

![Wilkinson 1990](Wilkinson1990_sm.png).

![Isaacdm](penalty_sm.png).

![Dennig](brexit_sm.png).

</div>

**Images:** (1) [Public license](https://commons.wikimedia.org/wiki/File:US_and_USSR_nuclear_stockpiles.svg), (2) [Wilkinson 1990](https://projects.ncsu.edu/project/evoresources/Evolutionary%20medicine/Unitsofselection%20DM%20copy/wilkinson-foodsharing.pdf), (3) [Isaacdm](https://commons.wikimedia.org/wiki/File:PhysicsofPenalty.JPG?uselang=en-gb), (4) [Public license](http://www.publicdomainpictures.net/view-image.php?image=180560&picture=brexit-referendum-uk)

## Conflict in conservation

Game theory provides a natural framework for addressing issues of cooperation and conflict in conservation.

- Game theory can be used to understand stake-holder decision making ([Kark et al. 2015 *Curr. Opin. Environ. Sustainability* **12**:12-24]()).

- Game-theoretic models can suggest novel management strategies ([Colvan et al. 2011 *Biol. Conserv.* **144**:1246-1253](http://dx.doi.org/10.1016/j.biocon.2010.10.028); [Tilman et al. 2016 *Theor. Ecol.*](http://link.springer.com/article/10.1007%2Fs12080-016-0318-8)).

- Game theory can consider multiple management objectives simultaneously ([Lee 2012 *Chemosphere* **87**:608-613](http://dx.doi.org/10.1016/j.chemosphere.2012.01.014)).

## Point about conflict and games being complex

Nils points out that typical game theory assumes that

- pay-offs are known and fixed
- Actions are clear and always available
- Players behave rationally
- Game rules are known to all
- There is a clear rank order of outcomes

## Genetic algorithms

Show the basic idea
Use the gamesGA package to illustrate it in a game-theoretic context

<!--- 
    START OF THE GENETIC ALGORITHM SET OF SLIDES  ------------------------------
--->

## Structure of a genetic algorithm

<div class="columns-2">

```{r, echo=FALSE}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}

par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,200), ylim=c(0,100), xaxt="n", yaxt="n",
     xlab="",ylab="", bty="n");
abox <- mbox(x0 = 0, x1 = 45, y0 = 70, y1 = 90);
polygon(x=abox$x, y=abox$y, lwd=3, border="red", col="rosybrown1");
text(x=22, y=85, labels="Random", col="black");
text(x=22, y=80, labels="seed of", col="black");
text(x=22, y=75, labels="strategies", col="black");
# --------------------------------------------------------------
cbox <- mbox(x0 = 70, x1 = 140, y0 = 0, y1 = 90);
polygon(x=cbox$x, y=cbox$y, lwd=3, border="goldenrod4", col="gold3");
arrows(x0=46, x1=70, y0=82, y1=82, lwd=2, length=0.15);
dbox <- mbox(x0 = 75, x1 = 135, y0 = 75, y1 = 88);
polygon(x=dbox$x, y=dbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=84, labels="Begin new", col="black");
text(x=105, y=79, labels="generation", col="black");
# ===============================================================
```

<br>

**Table of strategies**

```{r, echo=FALSE}
randvals <- sample(x=1:5, size=32, replace=TRUE);
agents   <- matrix(data=randvals, nrow=8);
rownames(agents) <- c("Strategy 1", "Strategy 2", "Strategy 3", "Strategy 4",
                      "Strategy 5", "Strategy 6", "Strategy 7", "Strategy 8");
colnames(agents) <- c("L1", "L2", "L3", "L4");
kable(agents, format="html");
```

</div>

## Structure of a genetic algorithm

<div class="columns-2">

```{r, echo=FALSE}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}

par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,200), ylim=c(0,100), xaxt="n", yaxt="n",
     xlab="",ylab="", bty="n");
abox <- mbox(x0 = 0, x1 = 45, y0 = 70, y1 = 90);
polygon(x=abox$x, y=abox$y, lwd=3, border="red", col="rosybrown1");
text(x=22, y=85, labels="Random", col="black");
text(x=22, y=80, labels="seed of", col="black");
text(x=22, y=75, labels="strategies", col="black");
# --------------------------------------------------------------
cbox <- mbox(x0 = 70, x1 = 140, y0 = 0, y1 = 90);
polygon(x=cbox$x, y=cbox$y, lwd=3, border="goldenrod4", col="gold3");
arrows(x0=46, x1=70, y0=82, y1=82, lwd=2, length=0.15);
dbox <- mbox(x0 = 75, x1 = 135, y0 = 75, y1 = 88);
polygon(x=dbox$x, y=dbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=84, labels="Begin new", col="black");
text(x=105, y=79, labels="generation", col="black");
# --------------------------------------------------------------
ebox <- mbox(x0 = 75, x1 = 135, y0 = 63, y1 = 73);
polygon(x=ebox$x, y=ebox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=68, labels="recombination", col="black");
# ===============================================================
```

<br>

**Table of strategies**

```{r, echo=FALSE}
for(i in 1:dim(agents)[1]){
    for(j in 1:dim(agents)[2]){
        check <- runif(n=1);
        if(check < 0.2){
            other            <- sample(x=1:dim(agents)[1], size=1);
            temp             <- agents[i,j];
            agents[i,j]      <- agents[other, j];
            agents[other, j] <- temp;
        }
    }
}
kable(agents, format="html");
```

</div>

## Structure of a genetic algorithm

<div class="columns-2">

```{r, echo=FALSE}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}

par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,200), ylim=c(0,100), xaxt="n", yaxt="n",
     xlab="",ylab="", bty="n");
abox <- mbox(x0 = 0, x1 = 45, y0 = 70, y1 = 90);
polygon(x=abox$x, y=abox$y, lwd=3, border="red", col="rosybrown1");
text(x=22, y=85, labels="Random", col="black");
text(x=22, y=80, labels="seed of", col="black");
text(x=22, y=75, labels="strategies", col="black");
# --------------------------------------------------------------
cbox <- mbox(x0 = 70, x1 = 140, y0 = 0, y1 = 90);
polygon(x=cbox$x, y=cbox$y, lwd=3, border="goldenrod4", col="gold3");
arrows(x0=46, x1=70, y0=82, y1=82, lwd=2, length=0.15);
dbox <- mbox(x0 = 75, x1 = 135, y0 = 75, y1 = 88);
polygon(x=dbox$x, y=dbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=84, labels="Begin new", col="black");
text(x=105, y=79, labels="generation", col="black");
# --------------------------------------------------------------
ebox <- mbox(x0 = 75, x1 = 135, y0 = 63, y1 = 73);
polygon(x=ebox$x, y=ebox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=68, labels="recombination", col="black");
# --------------------------------------------------------------
fbox <- mbox(x0 = 75, x1 = 135, y0 = 51, y1 = 61);
polygon(x=fbox$x, y=fbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=56, labels="mutation", col="black");
# ===============================================================
```

<br>

**Table of strategies**

```{r, echo=FALSE}
for(i in 1:dim(agents)[1]){
    for(j in 1:dim(agents)[2]){
        check <- runif(n=1);
        if(check < 0.2){
            agents[i,j]  <- sample(x=1:5, size=1);
        }
    }
}
kable(agents, format="html");
```

</div>

## Structure of a genetic algorithm

<div class="columns-2">

```{r, echo=FALSE}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}

par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,200), ylim=c(0,100), xaxt="n", yaxt="n",
     xlab="",ylab="", bty="n");
abox <- mbox(x0 = 0, x1 = 45, y0 = 70, y1 = 90);
polygon(x=abox$x, y=abox$y, lwd=3, border="red", col="rosybrown1");
text(x=22, y=85, labels="Random", col="black");
text(x=22, y=80, labels="seed of", col="black");
text(x=22, y=75, labels="strategies", col="black");
# --------------------------------------------------------------
cbox <- mbox(x0 = 70, x1 = 140, y0 = 0, y1 = 90);
polygon(x=cbox$x, y=cbox$y, lwd=3, border="goldenrod4", col="gold3");
arrows(x0=46, x1=70, y0=82, y1=82, lwd=2, length=0.15);
dbox <- mbox(x0 = 75, x1 = 135, y0 = 75, y1 = 88);
polygon(x=dbox$x, y=dbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=84, labels="Begin new", col="black");
text(x=105, y=79, labels="generation", col="black");
# --------------------------------------------------------------
ebox <- mbox(x0 = 75, x1 = 135, y0 = 63, y1 = 73);
polygon(x=ebox$x, y=ebox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=68, labels="recombination", col="black");
# --------------------------------------------------------------
fbox <- mbox(x0 = 75, x1 = 135, y0 = 51, y1 = 61);
polygon(x=fbox$x, y=fbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=56, labels="mutation", col="black");
# --------------------------------------------------------------
gbox <- mbox(x0 = 75, x1 = 135, y0 = 39, y1 = 49);
polygon(x=gbox$x, y=gbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=44, labels="check fitness", col="black");
# ===============================================================
```

<br>

**Table of strategies**

```{r, echo=FALSE}
fit <- apply(agents,1,sum);
kable(cbind(agents,fit), format="html");
```

</div>

## Structure of a genetic algorithm

<div class="columns-2">

```{r, echo=FALSE}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}

par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,200), ylim=c(0,100), xaxt="n", yaxt="n",
     xlab="",ylab="", bty="n");
abox <- mbox(x0 = 0, x1 = 45, y0 = 70, y1 = 90);
polygon(x=abox$x, y=abox$y, lwd=3, border="red", col="rosybrown1");
text(x=22, y=85, labels="Random", col="black");
text(x=22, y=80, labels="seed of", col="black");
text(x=22, y=75, labels="strategies", col="black");
# --------------------------------------------------------------
cbox <- mbox(x0 = 70, x1 = 140, y0 = 0, y1 = 90);
polygon(x=cbox$x, y=cbox$y, lwd=3, border="goldenrod4", col="gold3");
arrows(x0=46, x1=70, y0=82, y1=82, lwd=2, length=0.15);
dbox <- mbox(x0 = 75, x1 = 135, y0 = 75, y1 = 88);
polygon(x=dbox$x, y=dbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=84, labels="Begin new", col="black");
text(x=105, y=79, labels="generation", col="black");
# --------------------------------------------------------------
ebox <- mbox(x0 = 75, x1 = 135, y0 = 63, y1 = 73);
polygon(x=ebox$x, y=ebox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=68, labels="recombination", col="black");
# --------------------------------------------------------------
fbox <- mbox(x0 = 75, x1 = 135, y0 = 51, y1 = 61);
polygon(x=fbox$x, y=fbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=56, labels="mutation", col="black");
# --------------------------------------------------------------
gbox <- mbox(x0 = 75, x1 = 135, y0 = 39, y1 = 49);
polygon(x=gbox$x, y=gbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=44, labels="check fitness", col="black");
# --------------------------------------------------------------
hbox <- mbox(x0 = 75, x1 = 135, y0 = 27, y1 = 37);
polygon(x=hbox$x, y=hbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=32, labels="selection", col="black");
# ===============================================================
```

<br>

**Table of strategies**

```{r, echo=FALSE}
ord <- order(fit, decreasing=TRUE);
agents <- agents[ord,];
fit    <- fit[ord];
kable(cbind(agents,fit), format="html");
```

</div>

## Structure of a genetic algorithm

<div class="columns-2">

```{r, echo=FALSE}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}

par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,200), ylim=c(0,100), xaxt="n", yaxt="n",
     xlab="",ylab="", bty="n");
abox <- mbox(x0 = 0, x1 = 45, y0 = 70, y1 = 90);
polygon(x=abox$x, y=abox$y, lwd=3, border="red", col="rosybrown1");
text(x=22, y=85, labels="Random", col="black");
text(x=22, y=80, labels="seed of", col="black");
text(x=22, y=75, labels="strategies", col="black");
# --------------------------------------------------------------
cbox <- mbox(x0 = 70, x1 = 140, y0 = 0, y1 = 90);
polygon(x=cbox$x, y=cbox$y, lwd=3, border="goldenrod4", col="gold3");
arrows(x0=46, x1=70, y0=82, y1=82, lwd=2, length=0.15);
dbox <- mbox(x0 = 75, x1 = 135, y0 = 75, y1 = 88);
polygon(x=dbox$x, y=dbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=84, labels="Begin new", col="black");
text(x=105, y=79, labels="generation", col="black");
# --------------------------------------------------------------
ebox <- mbox(x0 = 75, x1 = 135, y0 = 63, y1 = 73);
polygon(x=ebox$x, y=ebox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=68, labels="recombination", col="black");
# --------------------------------------------------------------
fbox <- mbox(x0 = 75, x1 = 135, y0 = 51, y1 = 61);
polygon(x=fbox$x, y=fbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=56, labels="mutation", col="black");
# --------------------------------------------------------------
gbox <- mbox(x0 = 75, x1 = 135, y0 = 39, y1 = 49);
polygon(x=gbox$x, y=gbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=44, labels="check fitness", col="black");
# --------------------------------------------------------------
hbox <- mbox(x0 = 75, x1 = 135, y0 = 27, y1 = 37);
polygon(x=hbox$x, y=hbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=32, labels="selection", col="black");
# --------------------------------------------------------------
ibox <- mbox(x0 = 75, x1 = 135, y0 = 15, y1 = 25);
polygon(x=ibox$x, y=ibox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=20, labels="replacement", col="black");
# ===============================================================
```

<br>

**Table of strategies**

```{r, echo=FALSE}
agents <- rbind(agents[1,],
                agents[1,],
                agents[2,],
                agents[2,],
                agents[3,],
                agents[3,],
                agents[4,],
                agents[4,]
                );
fit    <- c(fit[1], fit[1], fit[2], fit[2], fit[3], fit[3], fit[4], fit[4]);
kable(cbind(agents,fit), format="html");
```

</div>

## Structure of a genetic algorithm

<div class="columns-2">

```{r, echo=FALSE}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}

par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,200), ylim=c(0,100), xaxt="n", yaxt="n",
     xlab="",ylab="", bty="n");
abox <- mbox(x0 = 0, x1 = 45, y0 = 70, y1 = 90);
polygon(x=abox$x, y=abox$y, lwd=3, border="red", col="rosybrown1");
text(x=22, y=85, labels="Random", col="black");
text(x=22, y=80, labels="seed of", col="black");
text(x=22, y=75, labels="strategies", col="black");
# --------------------------------------------------------------
cbox <- mbox(x0 = 70, x1 = 140, y0 = 0, y1 = 90);
polygon(x=cbox$x, y=cbox$y, lwd=3, border="goldenrod4", col="gold3");
arrows(x0=46, x1=70, y0=82, y1=82, lwd=2, length=0.15);
dbox <- mbox(x0 = 75, x1 = 135, y0 = 75, y1 = 88);
polygon(x=dbox$x, y=dbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=84, labels="Begin new", col="black");
text(x=105, y=79, labels="generation", col="black");
# --------------------------------------------------------------
ebox <- mbox(x0 = 75, x1 = 135, y0 = 63, y1 = 73);
polygon(x=ebox$x, y=ebox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=68, labels="recombination", col="black");
# --------------------------------------------------------------
fbox <- mbox(x0 = 75, x1 = 135, y0 = 51, y1 = 61);
polygon(x=fbox$x, y=fbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=56, labels="mutation", col="black");
# --------------------------------------------------------------
gbox <- mbox(x0 = 75, x1 = 135, y0 = 39, y1 = 49);
polygon(x=gbox$x, y=gbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=44, labels="check fitness", col="black");
# --------------------------------------------------------------
hbox <- mbox(x0 = 75, x1 = 135, y0 = 27, y1 = 37);
polygon(x=hbox$x, y=hbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=32, labels="selection", col="black");
# --------------------------------------------------------------
ibox <- mbox(x0 = 75, x1 = 135, y0 = 15, y1 = 25);
polygon(x=ibox$x, y=ibox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=20, labels="replacement", col="black");
# --------------------------------------------------------------
jbox <- mbox(x0 = 75, x1 = 135, y0 = 3, y1 = 13);
polygon(x=jbox$x, y=jbox$y, lwd=3, border="goldenrod4", col="white");
text(x=105, y=8, labels="converge?", col="black");
# --------------------------------------------------------------
bbox <- mbox(x0 = 0, x1 = 45, y0 = 20, y1 = 40);
polygon(x=bbox$x, y=bbox$y, lwd=3, border="blue", col="lightsteelblue1");
arrows(x0=70, x1=22.5, y0=5, y1=5, lwd=2, length=0.0);
arrows(x0=22, x1=22, y0=5, y1=20, lwd=2, length=0.15);
arrows(x0=55, x1=55, y0=5, y1=82, lwd=2, length=0.15);
text(x=15, y=10, labels="Yes", col="black", srt=90);
text(x=50, y=10, labels="No", col="black", srt=90);
text(x=22, y=33, labels="Adaptive", col="black");
text(x=22, y=28, labels="strategies", col="black");
# ===============================================================
```

<br>

**Table of strategies**

```{r, echo=FALSE}
kable(cbind(agents,fit), format="html");
```

</div>

<!--- 
    END OF THE GENETIC ALGORITHM SET OF SLIDES  --------------------------------
--->

##

![](mse_overview_sm.png)

##

![](gmse_overview_sm.png)

## Introduce GMSE

<div class = "columns-2">

![](GMSE_Logo_Goose_300.png)

- Objective of GMSE 1
- Objective of GMSE 2
- Objective of GMSE 3

</div>

## Introduce GMSE

<div class = "columns-2">

![](GMSE_Logo_Goose_300.png)

- Technical detail of GMSE 1
- Technical detail of GMSE 2
- Technical detail of GMSE 3

</div>

## Geese example

- Use some data from Jeremy's paper?

## Future software development