### 2018-04-24 notebook/ms/SI4.Rmd
New function to demonstrate

---
title: 'GMSE: an R package for generalised management strategy evaluation'
author: A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;,
  Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, O. Sarobidy Rakotonarivo&#xb9;,
  Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes: null
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
subtitle: Supporting Information 4
biblio-style: apalike
---

```{r, echo = FALSE}
library(GMSE);
```

Fine-tuning simulation conditions using `gmse_apply`
================================================================================

Here we demonstrate how simulations in GMSE can be more fine-tuned to specific empirical situations through the use of `gmse_apply`. To do this, we use the same scenario described in [Supporting Information 3](); we first recreate the basic scenario run in `gmse` using `gmse_apply`, and then build in additional modelling details including (1) [custom placement of user land](#land), (2) [parameterisation of individual user budgets](#budgets), and (3) [density-dependent movement of resources](#movement). We emphasise that these simulations are provided only to demonstrate the use of GMSE, and specifically to show the flexibility of the `gmse_apply` function, not to accurately recreate the dynamics of a specific system or make management recommendations.

We reconsider the case of a protected waterfowl population that exploits agricultural land [e.g., @Fox2017a; @Mason2017; @Tulloch2017; @Cusack2018]. The manager attempts to keep the watefowl at a target abundance, while users (farmers) attempt to maximise agricultural yield on the land that they own. We again parameterise our model using demographic information from the Taiga Bean Goose (*Anser fabalis fabalis*), as reported by @Johnson2018 and @AEWA2016. Relevant parameter values are listed in the table below.

| Parameter         | Value    | Description                                      |
|-------------------|----------|--------------------------------------------------|
| `remove_pr`       | 0.122    | Goose density-independent mortality probability  |
| `lambda`          | 0.275    | Expected offspring production per time step      |
| `res_death_K`     | 93870    | Goose carrying capacity (on adult mortality)     |
| `RESOURCE_ini`    | 35000    | Initial goose abundance                          |
| `manage_target`   | 70000    | Manager's target goose abundance                 |
| `res_death_type`  | 3        | Mortality (density and density-indepent sources) |

Table: GMSE simulation parameter values inspired by @Johnson2018 and @AEWA2016

Additionally, we continue to use the following values for consistency, except in the case of `stakeholders`, where we reduce the number of farmers to `stakeholders = 8` and increase `user_budget` to 100000. This is done to for two reasons. First, it speeds up simulations for the purpose of demonstration; second, it makes the presentation of landscape ownership easier (see below).

| Parameter         | Value    | Description                                      |
|-------------------|----------|--------------------------------------------------|
| `manager_budget`  | 10000    | Manager's budget for setting policy options      |
| `user_budget`     | 100000   | Users' budgets for actions                       |
| `public_land`     | 0.4      | Proportion of the landscape that is public       |
| `stakeholders`    | 8        | Number of stakeholders                           |
| `land_ownership`  | TRUE     | Users own landscape cells                        |
| `res_consume`     | 0.02     | Landscape cell output consumed by a resource     | 
| `observe_type`    | 3        | Observation model type (survey)                  |
| `agent_view`      | 1        | Cells managers can see when conducting a survey  |

Table: Non-default GMSE parameter values chosen by authors

All other values are set to GMSE defaults, except where specifically noted otherwise.

Re-creating `gmse` simulations using `gmse_apply`
================================================================================

We now recreate the simulations in [Supporting Information 3](), which were run using the `gmse` function, in `gmse_apply`. Doing so requires us to first initialise simulations using one run of `gmse_apply`, then loop through multiple time steps that again call `gmse_apply` and saving the results of interest. Following instructions in [Supporting Information 1](), we can call the initialising simulation `sim_old`, and use the code below to read in the relevant parameter values.

```{r, echo = TRUE, eval = TRUE}
sim_old  <- gmse_apply(get_res = "Full", remove_pr = 0.122, lambda = 0.275,
                       res_death_K = 93870, RESOURCE_ini = 35000, 
                       manage_target = 70000, res_death_type = 3,
                       manager_budget = 10000, user_budget = 100000,
                       public_land = 0.4, stakeholders = 8, 
                       land_ownership = TRUE, res_consume = 0.02, 
                       observe_type = 3, agent_view = 1);
```

Note that the argument `get_res = "Full"` causes `sim_old` retain all of the relevant data structures for simulating a new time step and recording simulation results. This includes the key simulation output, which is located in `sim_old$basic_output`, which is printed below.

```{r, echo = FALSE}
print(sim_old$basic_output);
```

We can then loop over 30 time steps to recreate the simulations from [Supporting Information 3](). In these simulations, we are specifically interested in the resource and observation outputs, as well as the manager policy and user actions for culling, which we record below in the data frame `sim_sum`. The inclusion of the argument `old_list` tells `gmse_apply` to use parameters and values from the list `sim_old` in the new time step.

```{r}
sim_sum_1 <- matrix(data = NA, nrow = 30, ncol = 5);
for(time_step in 1:30){
    sim_new                 <- gmse_apply(get_res = "Full", old_list = sim_old);
    sim_sum_1[time_step, 1] <- time_step;
    sim_sum_1[time_step, 2] <- sim_new$basic_output$resource_results[1];
    sim_sum_1[time_step, 3] <- sim_new$basic_output$observation_results[1];
    sim_sum_1[time_step, 4] <- sim_new$basic_output$manager_results[3];
    sim_sum_1[time_step, 5] <- sum(sim_new$basic_output$user_results[,3]);
    sim_old                 <- sim_new;
}
colnames(sim_sum_1) <- c("Time", "Pop_size", "Pop_est", "Cull_cost", 
                         "Cull_count");
print(sim_sum_1);
```

The above output from `sim_sum_1` shows the table that holds the information we were interested in pulling out of our simulation results. All of this information was available under the list element `sim_new$basic_output`, but other list elements of `sim_new` might also be useful to record. It is important to also remember that this example of `gmse_apply` is using the default resource, observation, manager, and user submodels. Custom submodels could produce different outputs in `sim_new` (see [Supporting Information 1]() for examples). For default options, there are some list elements that might be especially useful. All of these elements can be edited *within the above loop* to dynamically adjust simulations.

- `sim_new$resource_array`: A table holding all information on resources. Rows correspond to a unique resource, and columns correspond to resource properties: (1) ID, (2-4) resource types (not currently in use), (5) x-location, (6) y-location, (7) movement parameter, (8) time, (9) density independent mortality parameter (`remove_pr`), (10) reproduction parameter (`lambda`), (11) offspring number, (12) age, (13-14) observation columns, (15) consumption rate (`res_consume`), and (16-20) recorded experiences of user actions.
- `sim_new$AGENTS`: A table holding basic information on agents (manager and users). Rows correspond to a unique agent, and columns correspond to agent properties: (1) ID, (2) type (0 for the manager, 1 for users), (3-4) additional type options not currently in use, (5-6), x and y locations (usually ignored), (7) movement parameter (usually ignored), (8) time, (9) agent's viewing ability in cells (`agent_view`), (10) error parameter, (11-12) values for holding marks and tallies of resources, (13-15) values for holding observations, (16) yield from landscape cells, (17) budget (`manager_budget` and `user_budget`).
- `sim_new$observation_vector`: Estimate of total resource number from the observation model (`observation_array` also holds this information in a different way depending on `observe_type`)
- `sim_new$LAND`: The landscape on which interactions occur, which is stored as a 3D array with `land_dim_1` rows, `land_dim_2` columns, and 3 layers. Layer 1 (`sim_new$LAND[,,1]`) is not used, but could be used to store values that affect resources and agents. Layer 2 (`sim_new$LAND[,,2]`) stores crop yield from a cell, and layer 3 (`sim_new$LAND[,,3]`) stores the owner of the cell (value corresponds to the agent's ID).
- `sim_new$manage_vector`: The cost of each action as set by the manager. For even more fine-tuning, individual costs for actions can be set for each user in `sim_new$manager_array`.
- `sim_new$user_vector`: The total number of actions performed by each user. A more detailed breakdown of actions by individual users is held in `sim_new$user_array`.

We now show how to adjust the landscape to manually adjust land ownership in `gmse_apply`.

1. <a name="land">Custom placement of user land</a>
================================================================================

By default, all farmers in GMSE are allocated the same number of landscape cells, which are simply placed in order of the farmer's ID. Public landscape is produced by placing landscape cells that are technically owned by the manager, and therefore have landscape cell values of 0. The image below shows this landscape for the eight farmers from `sim_old`.

```{r}
image(x = sim_old$LAND[,,3], col = topo.colors(9), xaxt = "n", yaxt = "n");
```

We can change the ownership of cells by manipulating `sim_old$LAND[,,3]`. First we initialise a new `sim_old` below.

```{r, echo = TRUE, eval = TRUE}
sim_old  <- gmse_apply(get_res = "Full", remove_pr = 0.122, lambda = 0.275,
                       res_death_K = 93870, RESOURCE_ini = 35000, 
                       manage_target = 70000, res_death_type = 3,
                       manager_budget = 10000, user_budget = 100000,
                       public_land = 0.4, stakeholders = 8, 
                       land_ownership = TRUE, res_consume = 0.02, 
                       observe_type = 3, agent_view = 1);
```

Because we have not specified landscape dimensions in the above, the landscape reverts to the default size of 100 by 100 cells. We can then manually assign landscape cells to the eight farmers, whose IDs range from 2-9 (ID value 1 is the manager). Below we do this to make eight different sized farms.

```{r}
sim_old$LAND[1:20,   1:20,  3] <- 2; 
sim_old$LAND[1:20,  21:40,  3] <- 3;
sim_old$LAND[1:20,  41:60,  3] <- 4;
sim_old$LAND[1:20,  61:80,  3] <- 5;
sim_old$LAND[1:20,  81:100, 3] <- 6;
sim_old$LAND[21:40,  1:50,  3] <- 7;
sim_old$LAND[21:40, 51:100, 3] <- 8;
sim_old$LAND[41:60,  1:100, 3] <- 9;
sim_old$LAND[61:100, 1:100, 3] <- 1; # Public land
image(x = sim_old$LAND[,,3], col = topo.colors(9), xaxt = "n", yaxt = "n");
```

The above image shows the modified landscape stored in `sim_old`, which can now be incorporated into simulations using `gmse_apply`. We can think of all the plots on the left side of the landscape as farms of various sizes, while the blue area of the landscape on the right is public land.

2. <a name="budgets">Parameterisation of individual user budgets</a>
================================================================================

Perhaps we want to assume that farmers have different budgets, which are perhaps correlated in some way to the number of landscape cells that they own. Custom user budgets can be set by manipulating `sim_old$AGENTS`, the last column of which holds the budget for each user. Agent IDs (as stored on the landscape above) correspond to rows of `sim_old$AGENTS`, so individual budgets can be directly input as desired. We can do this manually (e.g., `sim_old$AGENTS[2, 17] <- 4000`), or, alternatively, if farmer budget positively correlates to landscape owned, we can use a loop to input values as below.

```{r}
for(ID in 2:9){
    cells_owned            <- sum(sim_old$LAND[,,3] == ID);
    sim_old$AGENTS[ID, 17] <- 100 * cells_owned;
}
```

The number of cells owned by each farmer is therefore listed in the table below.

```{r, echo = FALSE}
budget_table <- rbind(sim_old$AGENTS[,1], sim_old$AGENTS[,17]);
row.names(budget_table) <- c("ID", "Budget");
kable(budget_table, caption = "Custom created budgets for manager and users");
```

As with `sim_old$LAND` values, changes to `sim_old$AGENTS` will be retained in simulations looped through `gmse_apply`.

3. <a name="movement">Density-dependent movement of resources</a>
================================================================================

Lastly, we consider a more nuanced change to simulations, in which the rules for movement of resources are modified to account for density-dependence. Assume that geese tend to avoid aggregating, such that if a goose is located on the same cell as too many other geese, then it will move with a high probability at the end of a time step. Programming this movement rule can be accomplished by creating a new function to apply to the resource data array `sim_old$resource_array`. Below, a custom function is defined that causes a goose to move up to 5 cells in any direction if it finds itself on a cell with more than 10 other geese. As with default GMSE simulations, movement is based on a torus landscape.

```{r}
avoid_aggregation <- function(goose_table, land_dim_1 = 100, land_dim_2 = 100){
    goose_number <- dim(goose_table)[1]    # How many geese are there?
    for(goose in 1:goose_number){          # Loop through all rows of geese
        x_loc  <- goose_table[goose, 5];
        y_loc  <- goose_table[goose, 6];
        shared <- sum(goose_table[,5] == x_loc & goose_table[,6] == y_loc);
        if(shared > 10){
            new_x <- x_loc <- sample(x = -5:5, size = 1);
            new_y <- y_loc <- sample(x = -5:5, size = 1);
            if(new_x < 0){ # The 'if' statements below apply the torus
                new_x <- land_dim_1 + new_x;
            }
            if(new_x >= land_dim_1){
                new_x <- new_x - land_dim_1;
            }
            if(new_y < 0){
                new_y <- land_dim_2 + new_x;
            }
            if(new_y >= land_dim_2){
                new_y <- new_y - land_dim_2;
            }
            goose_table[goose, 5] <- new_x;
            goose_table[goose, 6] <- new_y;
        }
    }
    return(goose_table);
}
```