### 2019-01-04 notebook/presentation/GMSE_presentation.Rmd
Really working example

---
title: "Generalised Management Strategy Evaluation: modelling biodiversity and food security"
subtitle: "https://bradduthie.shinyapps.io/GMSE_presentation/#1"
author: "Brad Duthie (@bradduthie), Jeremy Cusack, Sarobidy Rakotonarivo, and Nils Bunnefeld"
date: "8 January 2018"
output: 
  ioslides_presentation:
    logo: Confoobio-stacked-logo.png
    css: slide.css
    widescreen: no
    fig_width: 5
    fig_height: 5
runtime: shiny
---

<style type="text/css">
body p {
  color: #000000;
}
ul {
  color: black;
}
ol {
  color: black;
}
a {
    color: darkblue;
    text-decoration: none
}
.fifty {
   font-size: 50%;
   }
.seventy {
   font-size: 70%;
   }
slides > slide.backdrop {
  background: none !important;
  background-color: white !important;
}
slides > slide.title-slide hgroup h1 {
  color: black;
}
slides > slide.title-slide p {
  color: black;
}
slides > slide.light > hgroup h2 {
  color: black;
}
slides > slide.light > hgroup h1 {
  color: black;
}
ul {
  color: black;
}
ol {
  color: black;
}
p {
  color: black
}
h1, h2, h3, h4, h5, h6, h7, h8, p {
    color: black;
}
.checkbox-inline, .radio-inline {
    margin-right: 90px;
}

.forceBreak { -webkit-column-break-after: always; break-after: column; }
</style>

```{r, echo=FALSE}
library(GMSE);
library(knitr);
library(shiny);
```

<!---
1. Know what to say about the Welikia project
--->

<!---
https://welikia.org/
http://ngm.nationalgeographic.com/2009/09/manhattan/miller-text
https://welikia.org/explore/mannahatta-map/
--->

# Biodiversity and food security

## {.smaller}

<center>

![](mannahatta_before.png)

</center>

## {.smaller}

<center>

![](mannahatta.png)

[Markley Boyer](https://www.facebook.com/markley.boyer), [Robert Clark](http://www.nationalgeographic.com/contributors/c/photographer-robert-clark/), [National Geographic](http://ngm.nationalgeographic.com/2009/09/manhattan/miller-text), [Welikia Project](https://welikia.org/)

</center>

## {.smaller}

<center>

![](potato_fields_Janet_Tench.jpg)

[Janet Tench](https://commons.wikimedia.org/wiki/File:Norfolk_potato_field_-_geograph.org.uk_-_180521.jpg)</footnote>

</center>

## {.smaller}

<center>

![](KehoeFig1.png)

[Kehoe et al. (2017) *Nat. Ecol. Evol.* **1**:1129–1135](https://www.nature.com/articles/s41559-017-0234-3)</footnote>

</center>

##

![](food_biodiversity.png)

##

![](wicked_modelling.png)

# Modelling biodiversity dynamics and human decision-making

##

<center>

![](mse_overview_sm.png)

</center>

## {.smaller}

<center>

![](mse_overview_sm_pt.png)

**Management Strategy Evaluation**

</center>

## {.smaller}

<center>

![](mse_overview_sm_un.png)

</center>
<br>

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; Bunnefeld et al. (2011) *Trends. Ecol. Evol.* [**26**:441-447](https://doi.org/10.1016/j.tree.2011.05.003) <br>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; Chadès et al. (2017) *Theor. Ecol.* [**10**:1-10](https://doi.org/10.1007/s12080-016-0313-0)

## The ConFooBio Project

**Resolving conflicts between food security and biodiversity conservation under uncertainty**

<div class = "columns-2">

![](confoobio_sans.png)

![](group.png)

</div>

## Generalised Management Strategy Evaluation (GMSE) software

<div class = "columns-2">

- **Model biodiversity dynamics & realistic human decision-making**
- **Predict resource & land-use changes in social-ecological systems**
- **Integrate flexibly with existing ecological models**
- **Open-source on [CRAN](https://cran.r-project.org/package=GMSE) and [GitHub](https://github.com/ConFooBio/gmse)**

<p class="forceBreak"></p>

<center>

![](GMSE_Logo_Goose_300.png)

[![](http://www.r-pkg.org/badges/version/GMSE?color=yellowgreen)](https://cran.r-project.org/package=GMSE) [![](http://cranlogs.r-pkg.org/badges/grand-total/GMSE?color=yellowgreen)](http://cranlogs.r-pkg.org/badges/grand-total/GMSE)

</center>

</div>
<br><br>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; Duthie AB, et al. (2018) *Method. Ecol. Evol.* [**9**:2396-2401](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.13091)

## GMSE: population modelling (defaults)

<div class = "columns-2"><br><br>

- **Default models individual-based (IBM)**
- **Spatially explicit landscape**
- **Individual birth, movement, death**
- **Can be over-ridden with custom models**

<p class="forceBreak"></p>
<br><br>
<center>

![](GMSE_Logo_Goose_300.png)

[![](http://www.r-pkg.org/badges/version/GMSE?color=yellowgreen)](https://cran.r-project.org/package=GMSE) [![](http://cranlogs.r-pkg.org/badges/grand-total/GMSE?color=yellowgreen)](http://cranlogs.r-pkg.org/badges/grand-total/GMSE)

</center>

</div>

## {.smaller}

<center>

![](case_studies.png)

<br>

Cusack, JJ, et al. *Conserv. Lett.* (*In press*). DOI: [10.1111/conl.12450](https://onlinelibrary.wiley.com/doi/full/10.1111/conl.12450)

</center>

##

<center>

![](mse_overview_sm.png)

</center>

##

<center>

![](gmse_overview_sm.png)

</center>

## Game theory: conflict and cooperation

<div class="columns-2">

- Formalisation and study of conflict and cooperation between rational decision-makers
- Compatibile with agent-based modelling
- Can suggest management strategiesand be used to understand stake-holder actions
- Conceptual unification

<p class="forceBreak"></p>

![](payoff_matrix_sm.png)
<center>
*Prisoner's dilemma*
</center>

</div><br>

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; Adami et al. (2016) *Phys. Life. Rev.* [**19**:1-26](http://dx.doi.org/10.1016/j.plrev.2016.08.015) <br>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; Tilman et al. (2016) *Theor. Ecol.* [**10**:155-165](http://link.springer.com/article/10.1007%2Fs12080-016-0318-8)

## {.smaller}

<div class="columns-2">

![Public license<footnote>test</footnote>](US_USSR_nukes_sm.png).

![Wilkinson 1990](Wilkinson1990_sm.png).

![Isaacdm](penalty_sm.png).

![Dennig](brexit_sm.png).

</div>

<br>

&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp; **Images:** (1) [Public license](https://commons.wikimedia.org/wiki/File:US_and_USSR_nuclear_stockpiles.svg), (2) [Wilkinson 1990](https://projects.ncsu.edu/project/evoresources/Evolutionary%20medicine/Unitsofselection%20DM%20copy/wilkinson-foodsharing.pdf), (3) [Isaacdm](https://commons.wikimedia.org/wiki/File:PhysicsofPenalty.JPG?uselang=en-gb), (4) [Public license](http://www.publicdomainpictures.net/view-image.php?image=180560&picture=brexit-referendum-uk)

## Game theory for wicked problems

<div class="columns-2">

<u>**Large strategy space:**</u>

 - **Asymmetric payoffs** (payoffs affected by stake-holder identity)
 - **Multiple choices**
 - **Multiple players**
 - **Iterated interactions** (strategies affected by interaction history)

<p class="forceBreak"></p>

![](payoff_matrix_sm.png)

## Game theory for wicked problems

<div class="columns-2">

<u>**Large strategy space:**</u>

 - **Asymmetric payoffs** (payoffs affected by stake-holder identity)
 - **Multiple choices**
 - **Multiple players**
 - **Iterated interactions** (strategies affected by interaction history)

<p class="forceBreak"></p>

![](magnus.jpg)
<center>
[Andreas Kontokanis](https://commons.wikimedia.org/wiki/File:Carlsen_Magnus_(30238051906).jpg)</footnote>
</center>

##

<center>

![](gmse_overview_sm.png)

</center>

## {.smaller}

<center>

![](gmse_overview_sm_ga2.png)

</center>

## GMSE: modelling decision-making

<div class = "columns-2">

**Up to 5 possible actions**

- Culling (&darr; survival)
- Castration (&darr; reproduction)
- Scaring (displacement)
- Feeding (&uarr; survival)
- Helping (&uarr; reproduction)

**Manager sets policy**

- Has a target population size
- Has a budget to set costs

<p class="forceBreak"></p>
<br><br>
<center>

![](GMSE_Logo_Goose_300.png)

[![](http://www.r-pkg.org/badges/version/GMSE?color=yellowgreen)](https://cran.r-project.org/package=GMSE) [![](http://cranlogs.r-pkg.org/badges/grand-total/GMSE?color=yellowgreen)](http://cranlogs.r-pkg.org/badges/grand-total/GMSE)

</center>

</div>

## GMSE: modelling decision-making

<div class = "columns-2">

**Up to 5 possible actions**

- Culling (&darr; survival)
- Castration (&darr; reproduction)
- Scaring (displacement)
- Feeding (&uarr; survival)
- Helping (&uarr; reproduction)

**Users respond to policy**

- Maximise land yield
- Maximise resource use

<p class="forceBreak"></p>

<center>

```{r, echo = FALSE}
sim_old  <- gmse_apply(get_res = "Full", remove_pr = 0.122, lambda = 0.275,
                       res_d4eath_K = 93870, RESOURCE_ini = 35000, 
                       manage_target = 70000, res_death_type = 3,
                       manager_budget = 10000, user_budget = 10000,
                       public_land = 0.4, stakeholders = 8, res_consume = 0.02, 
                       res_birth_K = 200000, land_ownership = TRUE,
                       observe_type = 3, agent_view = 1, converge_crit = 0.01,
                       ga_mingen = 200);
sim_old$LAND[1:20,   1:20,  3] <- 2; 
sim_old$LAND[1:20,  21:40,  3] <- 3;
sim_old$LAND[1:20,  41:60,  3] <- 4;
sim_old$LAND[1:20,  61:80,  3] <- 5;
sim_old$LAND[1:20,  81:100, 3] <- 6;
sim_old$LAND[21:40,  1:50,  3] <- 7;
sim_old$LAND[21:40, 51:100, 3] <- 8;
sim_old$LAND[41:60,  1:100, 3] <- 9;
sim_old$LAND[61:100, 1:100, 3] <- 1; # Public land
image(x = sim_old$LAND[,,3], col = topo.colors(9), xaxt = "n", yaxt = "n");
```

</center>

</div>

## GMSE: modelling decision-making {.smaller}

Default individual-based models can be run using `gmse` in R ([multiple options](https://confoobio.github.io/gmse/reference/gmse.html)).

```{r, eval = FALSE}
sim <- gmse(time_max = 30, land_dim_1 = 100, land_dim_2 = 100,
            observe_type = 1, res_death_K = 4000, 
            manager_budget = 2000, user_budget = 500, 
            scaring = TRUE, land_ownership = TRUE,
            tend_crops = TRUE, stakeholders = 7,
            res_consume = 1, RESOURCE_ini = 600);
```

```{r, eval = FALSE}
plot_gmse_results(sim);
```

```{r, eval = FALSE}
plot_gmse_effort(sim);
```

## {.smaller}

```{r, echo = FALSE}
plot_gmse_effort <- function(sim){

    agents <- sim$agents;
    paras  <- sim$paras;
    ACTION <- sim$action;
    COST   <- sim$cost;

    cols      <- c("green", "indianred1", "indianred3", "deepskyblue1",
                   "deepskyblue2");

    users    <- dim(agents[[1]])[1];
    max_time <- dim(paras)[1];

    para_vec <- paras[1,];

    allowed  <- sum(para_vec[89:93]);
    scar_cst  <- matrix(data = 0, nrow = max_time, ncol = users);
    cull_cst  <- matrix(data = 0, nrow = max_time, ncol = users);
    cast_cst  <- matrix(data = 0, nrow = max_time, ncol = users);
    feed_cst  <- matrix(data = 0, nrow = max_time, ncol = users);
    help_cst  <- matrix(data = 0, nrow = max_time, ncol = users);

    scar_act  <- matrix(data = 0, nrow = max_time, ncol = users);
    cull_act  <- matrix(data = 0, nrow = max_time, ncol = users);
    cast_act  <- matrix(data = 0, nrow = max_time, ncol = users);
    feed_act  <- matrix(data = 0, nrow = max_time, ncol = users);
    help_act  <- matrix(data = 0, nrow = max_time, ncol = users);
    scar_eff  <- matrix(data = 0, nrow = max_time, ncol = users);
    cull_eff  <- matrix(data = 0, nrow = max_time, ncol = users);
    cast_eff  <- matrix(data = 0, nrow = max_time, ncol = users);
    feed_eff  <- matrix(data = 0, nrow = max_time, ncol = users);
    help_eff  <- matrix(data = 0, nrow = max_time, ncol = users);    
    act_costs  <- matrix(data = 0, nrow = max_time, ncol = 5);
    pol_effort <- matrix(data = 0, nrow = max_time, ncol = 5);
    min_cost   <- para_vec[97];

    for(time in 1:max_time){
        #-- Scaring cost and actions
        scar_cst[time, 1] <- COST[[time]][3, 8, 1];
        for(user in 2:users){
            scar_cst[time, user] <- COST[[time]][1, 8, user];
        }
        scar_act[time, 1] <- ACTION[[time]][3, 8, 1] - min_cost;
        for(user in 2:users){
            scar_act[time, user] <- ACTION[[time]][1, 8, user];
        }
        scar_eff[time,] <- scar_act[time,]*scar_cst[time,]/agents[[time]][,17];
        #-- Culling cost and actions
        cull_cst[time, 1] <- COST[[time]][3, 9, 1];
        for(user in 2:users){
            cull_cst[time, user] <- COST[[time]][1, 9, user];
        }
        cull_act[time, 1] <- ACTION[[time]][3, 9, 1] - min_cost;
        for(user in 2:users){
            cull_act[time, user] <- ACTION[[time]][1, 9, user];
        }
        cull_eff[time,] <- cull_act[time,]*cull_cst[time,]/agents[[time]][,17];
        #-- Castration cost and actions
        cast_cst[time, 1] <- COST[[time]][3, 10, 1];
        for(user in 2:users){
            cast_cst[time, user] <- COST[[time]][1, 10, user];
        }
        cast_act[time, 1] <- ACTION[[time]][3, 10, 1] - min_cost;
        for(user in 2:users){
            cast_act[time, user] <- ACTION[[time]][1, 10, user];
        }
        cast_eff[time,] <- cast_act[time,]*cast_cst[time,]/agents[[time]][,17];
        #-- Feeding cost and actions
        feed_cst[time, 1] <- COST[[time]][3, 11, 1];
        for(user in 2:users){
            feed_cst[time, user] <- COST[[time]][1, 11, user];
        }
        feed_act[time, 1] <- ACTION[[time]][3, 11, 1] - min_cost;
        for(user in 2:users){
            feed_act[time, user] <- ACTION[[time]][1, 11, user];
        }
        feed_eff[time,] <- feed_act[time,]*feed_cst[time,]/agents[[time]][,17];
        #-- Helping cost and actions
        help_cst[time, 1] <- COST[[time]][3, 12, 1];
        for(user in 2:users){
            help_cst[time, user] <- COST[[time]][1, 12, user];
        }
        help_act[time, 1] <- ACTION[[time]][3, 12, 1] - min_cost;
        for(user in 2:users){
            help_act[time, user] <- ACTION[[time]][1, 12, user];
        }
        help_eff[time,] <- help_act[time,]*help_cst[time,]/agents[[time]][,17];
    }
    # -- Turn these into percentages:
    scar_eff <- scar_eff * 100;
    cull_eff <- cull_eff * 100;
    cast_eff <- cast_eff * 100;
    feed_eff <- feed_eff * 100;
    help_eff <- help_eff * 100;

    par(mfrow = c(5, 1), mar = c(0, 0, 0, 0), oma = c(6, 6, 0, 6));
    #---- Scaring
    if(para_vec[89] == 1){
        y1 <- 100;
        y2 <- y1 * 1.2
        plot(x = 1:max_time, y = 100 - scar_eff[,1], type = "l", lwd = 2, 
             cex.axis = 1.5, xaxt = "n", ylim = c(0, y2), yaxt = "n");
        axis(side = 2, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
        par(new = TRUE);
        plot(x = 1:max_time, y = scar_eff[,2], type = "n", lwd = 2, 
             ylim = c(0, y2), xaxt="n", yaxt="n", 
             cex.axis = 1.5);
        for(stakeholder in 2:users){
            points(x = 1:max_time, y = scar_eff[,stakeholder], type = "l", 
                   lwd = 1, col = cols[1]);
        }
        axis(side = 4, at = c(50, 100),  labels = c(50, 100), cex.axis = 1.5);
        legend(x = 1, y = y2 + y2*0.1, fill = cols[1:5], horiz = TRUE,
               legend = c("scaring", "culling", "castration", "feeding", 
                          "helping"), bty = "n", cex = 1);
        abline(h = y1 + y1*0.1, lwd = 2);
    }else{
        plot(x = 1:max_time, y = scar_eff[,1], type = "n", xaxt= "n", 
             yaxt = "n", ylim = c(0, 110));
        text(x = 0, y = 90, cex = 2, labels = "No scaring allowed",
             pos = 4);
        abline(h = 100, lwd = 2);
        legend(x = 1, y = 119, fill = cols[1:5], horiz = TRUE,
               legend = c("scaring", "culling", "castration", "feeding", 
                          "helping"), bty = "n", cex = 1);
    }
    #---- Culling
    if(para_vec[90] == 1){
        plot(x = 1:max_time, y = 100 - cull_eff[,1], type = "l", lwd = 2, 
             cex.axis = 1.5, xaxt = "n", ylim = c(0, y1), yaxt = "n");
        axis(side = 2, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
        par(new = TRUE);
        plot(x = 1:max_time, y = cull_eff[,2], type = "n", lwd = 2, 
             ylim = c(0, y1), xaxt="n", yaxt="n", cex.axis = 1.5);
        for(stakeholder in 2:users){
            points(x = 1:max_time, y = cull_eff[,stakeholder], type = "l", 
                   lwd = 1, col = cols[2]);
        }
        axis(side = 4, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
    }else{
        plot(x = 1:max_time, y = cull_act[,1], type = "n", xaxt= "n", 
             yaxt = "n", ylim = c(0, 100));
        text(x = 0, y = 90, cex = 2, labels = "No culling allowed",
             pos = 4);
    }
    #---- Castrating
    if(para_vec[91] == 1){
        plot(x = 1:max_time, y = 100 - cast_eff[,1], type = "l", lwd = 2, 
             cex.axis = 1.5, xaxt = "n", ylim = c(0, y1), yaxt = "n");
        axis(side = 2, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
        par(new = TRUE);
        plot(x = 1:max_time, y = cast_eff[,2], type = "n", lwd = 2, 
             ylim = c(0, y1), xaxt="n", yaxt="n", cex.axis = 1.5);
        for(stakeholder in 2:users){
            points(x = 1:max_time, y = cast_eff[,stakeholder], type = "l", 
                   lwd = 1, col = cols[3]);
        }
        axis(side = 4, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
    }else{
        plot(x = 1:max_time, y = cast_act[,1], type = "n", xaxt= "n", 
             yaxt = "n", ylim = c(0, 100));
        text(x = 0, y = 90, cex = 2, labels = "No castration allowed",
             pos = 4);
    }
    #---- Axes labels
    mtext("Manager's permissiveness of user action", 
          side = 2, line = 3.5, cex = 1.5, col = "black");
    mtext("Percentage of user action effort expended", 
          side = 4, line = 3.5, cex = 1.5, col = "black");
    #---- Feeding
    if(para_vec[92] == 1){
        plot(x = 1:max_time, y = 100 - feed_eff[,1], type = "l", lwd = 2, 
             cex.axis = 1.5, xaxt = "n", ylim = c(0, y1), yaxt = "n");
        axis(side = 2, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
        par(new = TRUE);
        plot(x = 1:max_time, y = feed_eff[,2], type = "n", lwd = 2, 
             ylim = c(0, y1), xaxt="n", yaxt="n", cex.axis = 1.5);
        for(stakeholder in 2:users){
            points(x = 1:max_time, y = feed_eff[,stakeholder], type = "l", 
                   lwd = 1, col = cols[4]);
        }
        axis(side = 4, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
    }else{
        plot(x = 1:max_time, y = feed_act[,1], type = "n", xaxt= "n", 
             yaxt = "n", ylim = c(0, 100));
        text(x = 0, y = 90, cex = 2, labels = "No feeding allowed",
             pos = 4);
    }
    #---- Helping
    if(para_vec[93] == 1){
        plot(x = 1:max_time, y = 100 - help_eff[,1], type = "l", lwd = 2, 
             cex.axis = 1.5, xaxt = "n", ylim = c(0, y1), yaxt = "n");
        axis(side = 2, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
        par(new = TRUE);
        plot(x = 1:max_time, y = help_eff[,2], type = "n", lwd = 2, 
             ylim = c(0, y1), xaxt="n", yaxt="n", cex.axis = 1.5);
        for(stakeholder in 2:users){
            points(x = 1:max_time, y = help_eff[,stakeholder], type = "l", 
                   lwd = 1, col = cols[5]);
        }
        axis(side = 4, at = c(50, 100), labels = c(50, 100), cex.axis = 1.5);
    }else{
        plot(x = 1:max_time, y = help_act[,1], type = "n", cex.axis = 1.5,
             yaxt = "n", ylim = c(0, 100));
        text(x = 0, y = 90, cex = 2, labels = "No helping offspring allowed",
             pos = 4);
    }    
    mtext("Time step", side = 1, line = 3.5, cex = 1.5, col = "black");
}
```

```{r, echo = FALSE}
inputPanel(
  actionButton("run", "Run Simulation"),
    actionButton("res", "Reset"),
  radioButtons("typ", "",
               c("General" = 1,
                 "Effort"  = 2),
               inline = TRUE, width = '800px')
)

sim      <- reactiveValues();
okay     <- reactiveValues();
sim$data <- NULL;
okay$go  <- 0;

sim_run <- observeEvent(input$run, {
              sim$data <- gmse(time_max = 30, land_dim_1 = 100, land_dim_2 = 100, 
                            observe_type = 1, res_death_K = 4000, plotting = FALSE,
                            manager_budget = 2000, user_budget = 500, 
                            scaring = TRUE, land_ownership = TRUE,
                            tend_crops = TRUE, stakeholders = 7,
                            res_consume = 1, RESOURCE_ini = 600)
              showNotification("Simulation has finished")
})

observeEvent(input$res, {
    sim$data <- NULL;
})

renderPlot(height = 500, {
    if(is.null(sim$data) == FALSE){
        if(input$typ == 1){
            plot_gmse_results(sim$data)
        }else{
            plot_gmse_effort(sim$data)
        }
    }
})
```

## {.smaller}

<br>
<center>

![](acknowledgements.png)

</center>

## {.smaller}

<br>
<center>

![](funding.png)

</center>