### 2022-04-14 notebook/cran_vignettes/SI9.Rmd
Initialise a new vignette to explain the simulated annealing approach

---
title: 'The simulated annealing algorithm of GMSE'
subtitle: 'GMSE: an R package for generalised management strategy evaluation (New function vignette)'
author: | 
        | Brad Duthie&#xb9; &#xb2;, Gabriela Ochoa&#xb9;
date: "[1] Biological and Environmental Sciences, University of Stirling, Stirling, UK [2] alexander.duthieh@stir.ac.uk"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes:
- \usepackage{lineno}
linestretch: 1
bibliography: gmse.bib
biblio-style: apalike
link-citations: true
linkcolor: blue
---

```{r, echo = FALSE}
library(GMSE);
# sim <- gmse(action_thres = 0.1, budget_bonus = 0.5, plotting = FALSE);
```

Overview of simulated annealing
================================================================================

The primary algorithm used in GMSE to model manager and user decision-making is the [genetic algorithm](SI1.pdf). As of version 1.0, GMSE also allows for the use of a simulated annealing algorithm [@Kirkpatrick1983; @Cerny1985; @Rutenbar1989]. Simulated annealing searches a space of potential solutions (in GMSE, decisions made by agents to maximise their own utility) in an iterative way. Unlike the [genetic algorithm](SI1.pdf), it is not a population focused approach.

Example use of simulated annealing
================================================================================

Performance of simulated annealing
================================================================================

References
================================================================================